--- 
title: "Gaming under lockdown: Steam data"
author: "Matti Vuorre"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: []
biblio-style: apalike
link-citations: yes
description: ""
---

# Preface

We study changes in gaming behavior during COVID-19 using data from Steam.

- GitHub repository: <https://github.com/digital-wellbeing/steam-lockdown>
- Website: <https://digital-wellbeing.github.io/steam-lockdown>
- OSF: <https://osf.io/ya9jt/>

## Analyses and reproducibility

The data analyses are organized into separate [R Markdown](https://rmarkdown.rstudio.com/) files for processing, describing, and modelling.

The project is organized as a R [bookdown](https://bookdown.org/yihui/bookdown/) project, so you can reproduce all analyses by building the book (e.g. in RStudio click the "Build Book" button). The results are rendered to `docs/index.html` and can be viewed in a web browser.

## Raw data

Is in data-raw/ which can't yet be pushed online.

## Reproducibility

```{r, include = FALSE}
# Ensure that .nojekyll exists for github pages
dir.create('docs', F); file.create('docs/.nojekyll')
```

```{r}
options(width = 120)
library(sessioninfo)
session_info()
```
